Analysis

### 2022-11-14 04_FIMO_processing.Rmd
CTCF Overlap analysis

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
library(GenomicRanges)
library(rtracklayer)
```

@@

```{r}
# for file in `find . -type f -name "*MA0139*"`; do
#   echo $file
#   newfile=`echo $file | sed 's/MA0139/JASPAR2022_MA0139/'`
#   echo $newfile
#   mv $file $newfile
# done
# 
# Output files
fileNameOut1 <- file.path(dir_project, "results", "log_PWMs.csv")
if (file.exists(fileNameOut1)) { file.remove(fileNameOut1) }
# Defined from FIMO_EDA
pval_threshold <- 1e-6
files <- list.files(file.path(dir_project, "RData"), pattern = "bed", full.names = TRUE)
for (fil in files) {
  print(basename(fil) %>% str_remove(., ".bed"))
  mtx <- read_tsv(fil, col_names = FALSE, col_types = c("ciicdciddc"))
  colnames(mtx) <- c("seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence")
  gr <- makeGRangesFromDataFrame(mtx, keep.extra.columns = TRUE)
  # Summary
  summary_to_save <- data.frame(File = basename(fil) %>% str_remove(., ".bed"),
                                Assembly = strsplit(basename(fil), ".", fixed = TRUE)[[1]][1],
                                `All (p-value threshold 1e-4)` = length(gr),
                                `Reduced (p-value threshold 1e-4)` = length(reduce(gr)),
                                `All (p-value threshold 1e-6)` = length(gr[gr$pvalue < pval_threshold]),
                                `Reduced (p-value threshold 1e-6)` = length(reduce(gr[gr$pvalue < pval_threshold])) )
  write_csv(summary_to_save, fileNameOut1, append = TRUE)
  gc()
}
```

>>>>

```{r libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(cowplot)
library(stringr)
library("ggsci")
library(scales)
# scales::show_col(pal_lancet("lanonc")(8))
mycols = pal_lancet("lanonc")(8)
library(GenomicRanges)
library(rtracklayer)
library(GenomicDistributions)
```

@@

```{r}
# for file in `find . -type f -name "*MA0139*"`; do
#   echo $file
#   newfile=`echo $file | sed 's/MA0139/JASPAR2022_MA0139/'`
#   echo $newfile
#   mv $file $newfile
# done
# 
# Output files
fileNameOut1 <- file.path(dir_project, "results", "log_PWMs.csv")
if (file.exists(fileNameOut1)) { file.remove(fileNameOut1) }
# Defined from FIMO_EDA
pval_threshold <- 1e-6
files <- list.files(file.path(dir_project, "RData", "BEDfiles"), pattern = "bed", full.names = TRUE) %>% sort
for (fil in files) {
  print(basename(fil) %>% str_remove(., ".bed"))
  mtx <- read_tsv(fil, col_names = FALSE, col_types = c("ciicdciddc"))
  colnames(mtx) <- c("seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence")
  gr <- makeGRangesFromDataFrame(mtx, keep.extra.columns = TRUE)
  # Width distribution
  gr_width <- width(gr)
  width_distribution <- paste(formatC(min(gr_width), format = "f", digits = 0), 
                              formatC(mean(gr_width), format = "f", digits = 2), 
                              formatC(max(gr_width), format = "f", digits = 0), sep = "/")
  # Distance to nearby
  gr_nearest <- calcNearestNeighbors(gr)
  nearest_distribution <- paste(formatC(min(gr_nearest), format = "f", digits = 0), 
                              formatC(mean(gr_nearest), format = "f", digits = 2), 
                              formatC(max(gr_nearest), format = "f", digits = 0), sep = "/")
  # # Proportion of overlapping enrties
  # tmp <- findOverlaps(gr, gr)
  # prop_overlapping <- sort(table(queryHits(tmp)) %>% table(), decreasing = TRUE)
  # prop_overlapping <- sum(prop_overlapping[which(names(prop_overlapping) != "1")]) / length(gr)
  # Summary
  summary_to_save <- data.frame(File = basename(fil) %>% str_remove(., ".bed"),
                                Assembly = strsplit(basename(fil), ".", fixed = TRUE)[[1]][1],
                                `Width, bp (min/mean/max)` = width_distribution,
                                `Distance to nearest, bp (min/mean/max)` = nearest_distribution,
#                                 `Proportion of overlapping` = formatC(prop_overlapping, format = "f", digits = 3),
                                `All (p-value threshold 1e-4)` = length(gr),
                                `Reduced (p-value threshold 1e-4)` = length(reduce(gr)),
                                `All (p-value threshold 1e-6)` = length(gr[gr$pvalue < pval_threshold]),
                                `Reduced (p-value threshold 1e-6)` = length(reduce(gr[gr$pvalue < pval_threshold])) )
  write_csv(summary_to_save, fileNameOut1, append = TRUE)
  gc()
}
```

# Overlap

```{r}
files <- list.files(file.path(dir_project, "RData", "BEDfiles"), pattern = "bed", full.names = TRUE) %>% sort
# Vector to store proportion of overlapping sites
prop_overlapping <- vector(mode = "numeric", length = length(files))
for (i in 1:length(files)) {
  fil <- files[i]
  print(basename(fil) %>% str_remove(., ".bed"))
  # Process all but CIS_BP data
  if (!grepl("CIS_BP", fil)) {
    mtx <- read_tsv(fil, col_names = FALSE, col_types = c("ciicdciddc"))
    colnames(mtx) <- c("seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence")
    gr <- makeGRangesFromDataFrame(mtx, keep.extra.columns = TRUE)
    # Proportion of overlapping enrties
    tmp <- findOverlaps(gr, gr)
    prop_overlap <- sort(table(queryHits(tmp)) %>% table(), decreasing = TRUE)
    prop_overlapping[i] <- sum(prop_overlap[which(names(prop_overlap) != "1")]) / length(gr)
  } else {
    prop_overlapping[i] <- NA
  }
}
prop_overlap_df <- data.frame(Data = basename(files) %>% str_remove(., ".bed"),
                              Prop_overlap = prop_overlapping)
```
